Quality Assesment of OpenStreetMap Footpath Data for 3D City Modelling

### 2020-05-09 report/report.tex
Update figure references
In-text figure references

Other than for the purpose of mentioning various sources or error, this project will focus exclusively on the Positional accuracy, Completeness and Semantic accuracy classes. The comparison of the OSM and SLU maps will focus on these accuracy classes to determine the error tolerance when manipulating features in the 3D map presentation and the road network generating phase.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_compare}
    \caption{Example segments of the Stockholm metropolitan area with unsystematic errors in precision. The SLU building dataset (blue) in superimposed on the OSM dataset (dark gray). The top picture is centered on a city block which shows a semantic mismatching between the two datasets. The whole block consists of only two buildings in the OSM dataset, but is divided into 15 smaller building lots in the SLU dataset. The bottom picture shows 4 smaller buildings whose the positional accuracy is low, due to the buildings being incorrectly scaled, skewed and rotated.}
    \label{fig:space}
\end{figure}

@@

(Fan et al., 2014) used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data. The motivation for this is to eliminate semantic differences between both datasets. A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set. This makes one-to-one object mapping impossible. Thus using building area as an estimator for completeness solves this issue much better than i.e. the number of buildings or other objects in each dataset.
This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_building_match}
    \caption{Examples of a 1:1 building match (left), where the sought-after building was found both in the OSM (red) and SLU (blue) datasets, and of a 1:n building match (right), where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.}
    \label{fig:space}
\end{figure}

@@

(Fan et al., 2014) further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.
This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
Figure 1 under Section 5 shows a typical comparison between the OSM and SLU datasets.
By eyesight it was determined that the displacement of features is small.

@@

The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning funcitons of matching building footprints in both datasets.
The turning function was first defined by Arkin et al. (1991), as a method for measuring the similarity of two polygons. The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l. This project uses the turning function as it is defined by (Fan et al., 2014). For a polygon with vertices ${v_1 ... v_n}$ and line segments ${e_1 ... e_n}$ It is defined as follows.
Fix a starting vertex $v_1$.
The tangent angle at $v_1$ is $\theta_{n,1}$. This is the angle between the neighbouring line segments $e_n$ and $e_1$.
For any $i$ such that $i>1$ and $n<i$, the tangent angle at $v_i$ is recursively defined as:
\begin{center}
    $\theta_{i, i-1} = \theta_{n,1} + \sum^{i}_{k=1} \theta_{k, k-1}$
\end{center}
The turning function has some nice geometric properties, in that it is invariant to both rotation and scaling of the polygon. The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under roation. It also measures only the normalized cumulative length, which does not change under scaling.
The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions:
\begin{center}
    $S_{T}(A, B) = 1 - (\int^{1}_{0} T_{C,A}(l) - T_{C,B}(l) dl)^{1/2}$.
\end{center}
The value range will be $(0 < S_{T} < 1)$, where $S_{T}(A,B) = 1$ if the polygons are identical.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function}
    \caption{An example of the turning function of a polygon.}
    \label{fig:space}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function_diff_filled}
    \caption{An example of the turning functions of two corresponding buildings, with the area between them highlighted.}
    \label{fig:space}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_pos_error_plot}
    \caption{The average positional accuracies per building as a frequency diagram.}
    \label{fig:space}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turning_function_plot}
    \caption{The turning function distances between 1:1 building pairs as a frequency diagram.}
    \label{fig:space}
\end{figure}

@@

The low semantic accuracy has made analysis hard. The available SLU dataset from Lantmäteriet uses a plot-based subdivision of buildings, whereas OSM uses a subdivision that more closely resembles the actual building footprint as seen from above. In many cases whole city blocks are represented as single buildings, which means that many less meaningful feature matches can be found.
% TODO: Table reference
As seen in table ***, about 80\% of all OSM features were found to have a 1:1 mapping to the SLU dataset, but upon further visual comparison it can be seen that many of the 1:n matches are found in a very specified area, namely the residential areas in downtown, Stockholm. This is problematic since this typical area with the same level of accessibility was likely measured with the same measuring technique, and the unsystematic positional error will likely be similar in these areas. The question is wether it comforms well to the positional error in the rest of the dataset or not. This analysis will not be carried out in this project, but could be a good subject for a future master's thesis project.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_utility_buildings}
    \caption{A map segment which shows an area with many small utility buildings that are represented in the SLU dataset but not in the OSM dataset.}
    \label{fig:space}
\end{figure}

@@

Each case is exemplified in figure 8.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_cases}
    \caption{The 5 identified cases over feature overlap. Top left: Case 1. Top right: Case 2. Middle: Case 3. Bottom left: Case 4. Bottom right: Case 5.}
    \label{fig:space}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_1}
    \caption{Case 1 collisions can be solved by translating colliding way points away from the building, along the normal of the nearest surface. Smoothing can then be applied as seen fit.}
    \label{fig:space}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_2}
    \caption{Case 3 collisions can be solved by eliminating the building features that intersect with the way, and then applying the same algorithm as in case 1.}
    \label{fig:space}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_3}
    \caption{Case 4 collisions can be solved by calculating the average normal over all colliding edges for both roads, and translating colliding points along the normal, away from the other road. Smoothing can then be applied as seen fit.}
    \label{fig:space}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_4}
    \caption{Case 5 collisions can be solved simply by translating the colliding point along the tangent of the closest edge, and then applying smoothing as seen fit.}
    \label{fig:space}
\end{figure}

>>>>

Other than for the purpose of mentioning various sources or error, this project will focus exclusively on the Positional accuracy, Completeness and Semantic accuracy classes.
The comparison of the OSM and SLU maps will focus on these accuracy classes to determine the error tolerance when manipulating features in the 3D map presentation and the road network generating phase.
Figure \ref{fig:osm-slu-map} shows two segments of the Stockholm map where the OSM and SLU datasets show lower semantic and positional accuracy respectively.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_compare}
    \caption{Example segments of the Stockholm metropolitan area with unsystematic errors in precision. The SLU building dataset (blue) in superimposed on the OSM dataset (dark gray). The top picture is centered on a city block which shows a semantic mismatching between the two datasets. The whole block consists of only two buildings in the OSM dataset, but is divided into 15 smaller building lots in the SLU dataset. The bottom picture shows 4 smaller buildings whose the positional accuracy is low, due to the buildings being incorrectly scaled, skewed and rotated.}
    \label{fig:osm-slu-map}
\end{figure}

@@

(Fan et al., 2014) used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
The motivation for this is to eliminate semantic differences between both datasets.
A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set.
This makes a complete one-to-one object mapping impossible.
Rather, buildings have to be identified as 1:1 (one to one), 1:n (one to many) or 1:0 (one to none) cases, depending on if a building is fully represented in the other dataset, represented as an aggregation of smaller plots, or not found at all.
Figure \ref{fig:building-match-types} shows examples of two buildings that were found in the OSM and SLU datasets as 1:1 and 1:n matches respectively.

Since it is impossible to find a one-to-one feature map for building footprints, using building area as an estimator for completeness is much better than i.e. using the number of buildings or other objects in each dataset.
This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_building_match}
    \caption{Examples of a 1:1 building match (left), where the sought-after building was found both in the OSM (red) and SLU (blue) datasets, and of a 1:n building match (right), where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.}
    \label{fig:building-match-types}
\end{figure}

@@

(Fan et al., 2014) further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.
This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
Figure \ref{fig:osm-slu-map} under Section 5 shows a typical comparison between the OSM and SLU datasets.
By eyesight it was determined that the displacement of features is small.

@@

The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning funcitons of matching building footprints in both datasets.
The turning function was first defined by Arkin et al. (1991), as a method for measuring the similarity of two polygons.
The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
This project uses the turning function as it is defined by (Fan et al., 2014).
See figure \ref{fig:turning-function} for a side-by-side comparison of a polygon and its turning function.
For a polygon with vertices ${v_1 ... v_n}$ and line segments ${e_1 ... e_n}$ It is defined as follows.
Fix a starting vertex $v_1$.
The tangent angle at $v_1$ is $\theta_{n,1}$. This is the angle between the neighbouring line segments $e_n$ and $e_1$.
For any $i$ such that $i>1$ and $n<i$, the tangent angle at $v_i$ is recursively defined as:
\begin{center}
    $\theta_{i, i-1} = \theta_{n,1} + \sum^{i}_{k=1} \theta_{k, k-1}$
\end{center}
The turning function has some nice geometric properties, in that it is invariant to both rotation and scaling of the polygon. The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under roation. It also measures only the normalized cumulative length, which does not change under scaling.
The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions:
\begin{center}
    $S_{T}(A, B) = 1 - (\int^{1}_{0} T_{C,A}(l) - T_{C,B}(l) dl)^{1/2}$.
\end{center}
The value range will be $(0 < S_{T} < 1)$, where $S_{T}(A,B) = 1$ if the polygons are identical. 
See figure \ref{fig:turning-function-compare} for a side-by-side comparison of two similar polygons and their turning functions.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function}
    \caption{An example of the turning function of a polygon.}
    \label{fig:turning-function}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function_diff_filled}
    \caption{An example of the turning functions of two corresponding buildings from the OSM and SLU datasets, with the area between them highlighted.}
    \label{fig:turning-function-compare}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_pos_error_plot}
    \caption{The average positional accuracies per building as a frequency diagram.}
    \label{fig:bar-plot-positional-accuracy}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turning_function_plot}
    \caption{The shape similarities between 1:1 building pairs as a frequency diagram.}
    \label{fig:bar-plot-similarity}
\end{figure}

@@

The low semantic accuracy has made analysis hard.
The available SLU dataset from Lantmäteriet uses a plot-based subdivision of buildings, whereas OSM uses a subdivision that more closely resembles the actual building footprint as seen from above.
In many cases whole city blocks are represented as single buildings, which means that many less meaningful feature matches can be found.
As seen in table \ref{table:building-matching}, about 80\% of all OSM features were found to have a 1:1 mapping to the SLU dataset, but upon further visual comparison it can be seen that many of the 1:n matches are found in a very specified area, namely the residential areas in downtown, Stockholm.
This is problematic since this typical area with the same level of accessibility was likely measured with the same measuring technique, and the unsystematic positional error will likely be similar in these areas.
The question is wether it comforms well to the positional error in the rest of the dataset or not.
This analysis will not be carried out in this project, but could be a good subject for a future master's thesis project.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_utility_buildings}
    \caption{A map segment which shows an area with many small utility buildings that are represented in the SLU dataset but not in the OSM dataset.}
    \label{fig:osm-slu-map-utility-buildings}
\end{figure}

@@

Each case is exemplified in figure \ref{fig:collision-cases}.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_cases}
    \caption{The 5 identified cases over feature overlap. Top left: Case 1. Top right: Case 2. Middle: Case 3. Bottom left: Case 4. Bottom right: Case 5.}
    \label{fig:collision-cases}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_1}
    \caption{Case 1 collisions can be solved by translating colliding way points away from the building, along the normal of the nearest surface. Smoothing can then be applied as seen fit.}
    \label{fig:collision-case-1}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_2}
    \caption{Case 3 collisions can be solved by eliminating the building features that intersect with the way, and then applying the same algorithm as in case 1.}
    \label{fig:collision-case-3}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_3}
    \caption{Case 4 collisions can be solved by calculating the average normal over all colliding edges for both roads, and translating colliding points along the normal, away from the other road. Smoothing can then be applied as seen fit.}
    \label{fig:collision-case-4}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_4}
    \caption{Case 5 collisions can be solved simply by translating the colliding point along the tangent of the closest edge, and then applying smoothing as seen fit.}
    \label{fig:collision-case-5}
\end{figure}
